---
title: "Data Types"
format: html
---

```{r}
print_label <- function(label) {
  cat("\n")
  cat(label)
  cat("\n")
  cat(paste0(rep("=", nchar(label)), collapse = ""))
  cat("\n")
}
```


## Vectors

- Vectors are the basic data structure in R (atomic)
- All the elements in a vector are of the same type (homogeneous).  
- A vector can be though of as a single column with values of the same data type.
```{r}
# integer vector
(int_vec = c(1, 2, 3, 4))
```

```{r}
# character vector
(cha_vec = c("a", "b", "c", "d"))
```

## Lists

- Lists can hold different data types, including, vectors, other lists, and data frames (heterogeneous).

```{r}
(lst1 = list(1, 2, "x", "y"))
```

```{r}
(lst2 = list(c(1, 2), c("x", "y")))
```

```{r}
(list3 = list(c(1, 2, "x", "y")))
```

## Naming 

The elements vectors and lists can be named.

```{r}
(named_int_vec = c(a = 1, b = 2, c = 3, d = 4))
named_int_vec |> class()

print_label('named_int_vec[1]')
named_int_vec[1]
named_int_vec[1] |> class()

print_label('named_int_vec["a"]')
named_int_vec["a"]
named_int_vec["a"] |> class()

print_label('named_int_vec$1')
# named_int_vec$1

print_label('named_int_vec$a')
# named_int_vec$a

print_label('named_int_vec[1]')
named_int_vec[[1]]
named_int_vec[[1]] |> class()

print_label('named_int_vec["a"]')
named_int_vec[["a"]]
named_int_vec[["a"]] |> class()
```

```{r}
(named_lst1 = list(a = 1, b = 2, c = "x", d = "y"))
named_lst1 |> class()

print_label('named_lst1[1]')
named_lst1[1]
named_lst1[1] |> class()

print_label('named_lst1["a"]')
named_lst1["a"]
named_lst1["a"] |> class()

print_label('named_lst1$1')
# named_lst1$1

print_label('named_lst1$a')
named_lst1$a
named_lst1$a |> class()

print_label('named_lst1[[1]]')
named_lst1[[1]]
named_lst1[[1]] |> class()

print_label('named_lst1[["a"]]')
named_lst1[["a"]]
named_lst1[["a"]] |> class()
```

## Dataframes

Dataframes are special type of named lists where 

- values in each column are of the same type (homogeneous)
- columns are of the same length

```{r}
(df = data.frame(a = c(1, 2, 3), b = c("x", "y", "z")))
df |> class()

print_label('df[1]')
df[1]
df[1] |> class()

print_label('df["a"]')
df["a"]
df["a"] |> class()

print_label('df$1')
# df$1

print_label('df$a')
df$a
df$a |> class()

print_label('df[[1]]')
df[[1]]
df[[1]] |> class()

print_label('df[["a"]]')
df[["a"]]
df[["a"]] |> class()
```

